---
title: "Toronto Shelters"
subtitle: "Analyzing the State of Homeless Shelters in Toronto"
author: 
  - Ritvik Puri
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper presents a report on the conditions of various homeless shelters across the city of Toronto. Various factors such as location, occupancy, program type are taken into account. The aim is to find areas where these shelters are lacking and further imporvement cant be made."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(opendatatoronto)
library(dplyr)
```

```{r, include=FALSE}

# get package
package <- show_package("21c83b32-d5a8-4106-a54f-010dbe49f6f2")
package

# get all resources for this package
resources <- list_package_resources("21c83b32-d5a8-4106-a54f-010dbe49f6f2")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
view(data)
```

```{r, include = FALSE}
data %>%
  ggplot(aes(x = OCCUPIED_ROOMS, y = UNOCCUPIED_ROOMS)) +
  geom_point()

length(unique(data$SHELTER_ID))
```

# Introduction

Homelessness is a big problem in almost every country around the world. Despite being one of the most deveoped countries in the world, Canada has around quarter of a million people living wihtout a roof on their head. As you may expect, these rates are even higher for a busy metropolitan city such as Toronto. For every 10,000 people, 30 are homeless. While this may not sound like a lot at first glance, there are over 2 million people living in the city. This brings up mulitple questions. Who are these people? How do they live? And perhaps most importantly, what is the government doing for them?

The government of Canada has a lot of shelters for helping such people. The city of Toronto has 65 such shelters, each of various services and programs to help the people that they take in. This paper aims to analyze and report the conditions of these shelters, such as how much are they being used? Is a specific area in the city a hot spot for homeless people? Do certain programs get less people than others?

\newpage

# Data


# Results

# Discussion

## First discussion point

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


